---
title: "Allometry plus strains"
author: "susi"
date: "10/05/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## loading rds files

### allometry: 
based on 20191203-ControlsSexDimorphism.csv.gz  (52.7 MB) // 20191203-ControlsSexDimorphism.csv (390.1 MB; continous data only, 2.87 millian observations)

### strains: 
(contains also categorical etc data, over 10 million obs)
20191203-ControlsSexDimorphism.csv.gz  (108.1 MB) // 20191203-ControlsSexDimorphism.csv (1.7 GB)

Note: csv files were extrated from gz and loaded in R, saved as .rds WITHOUT manipulation (easier to handle / smaller size). 
Any additial data manipulation starts below

```{r}
#saveRDS(allometry, file= "allometry.rds")

allometry <- readRDS(file = "allometry.rds")
str(allometry)
```


```{r}
#saveRDS(strains, file= "strainsy.rds")

strains <- readRDS(file = "strains.rds")
str(strains)
```


```{r}
library(data.table)
library(table.express)

allometryDT <- setDT(allometry)
strainsDT <- setDT(strains[, c("external_sample_id", "strain_accession_id","strain_name")]) 

strains2 <- unique(strainsDT)
  
allometryNEW <-  merge(allometryDT,strains2,by="external_sample_id")
 
saveRDS(allometryNEW, file= "allometryNEW.rds")

```


